### 2020-10-04 cover-letter/cover-letter.Rmd
added cover letter

---
from: Earo Wang
fromqualifications: PhD (Statistics)
fromtitle: Lecturer in Statistics
fromWWW: earo.me
fromemail: earo.wang@auckland.ac.nz
fromphone: +64 9 923 9692
fromsignature: signature.png
toname: Editor
toaddress: The R Journal
opening: Dear Dr Michael Kane,
closing: Sincerely
spacing: 1.4
output: UOARmdTemplates::letter
---

Please consider our manuscript titled, "Conversations in time: interactive visualisation to explore structured temporal data" for publication in the R Journal.

This paper describes a new R package, **tsibbletalk**, for interactively exploring temporal data of nesting and crossing structures using linked views and time wrapping. It implements a shared tsibble data class to allow for linked brushing between html widgets on both client and server sides. It also provides a shiny module to wrap time lines for revealing periodic or aperiodic patterns.

We discuss **tsibbletalk** functions to easily build linked and dynamic graphics for temporal data, and include code examples using two contemporary data applications. The paper is written in RMarkdown, which contains the R code to reproduce the results.

We believe that this manuscript is a good fit for the R Journal, because it features novel implementation and nimble user experience in R.