### 2021-07-18 tsibbletalk.tex

% !TeX root = RJwrapper.tex
\title{Conversations in time: interactive visualization to explore
structured temporal data}
\author{by Earo Wang and Dianne Cook}

\maketitle

\abstract{%
Temporal data often has a hierarchical structure, defined by categorical
variables describing different levels, such as political regions or
sales products. The nesting of categorical variables produces a
hierarchical structure. The \CRANpkg{tsibbletalk} package is developed
to allow a user to interactively explore temporal data, relative to the
nested or crossed structures. It can help to discover differences
between category levels, and uncover interesting periodic or aperiodic
slices. The package implements a shared \texttt{tsibble} object that
allows for linked brushing between coordinated views, and a shiny module
that aids in wrapping timelines for seasonal patterns. The tools are
demonstrated using two data examples: domestic tourism in Australia and
pedestrian traffic in Melbourne.
}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Temporal data typically arrives as a set of many observational units
measured over time. Some variables may be categorical, containing a
hierarchy in the collection process, that may be measurements taken in
different geographic regions, or types of products sold by one company.
Exploring these multiple features can be daunting. Ensemble graphics
\citep{unwin_ensemble_2018} bundle multiple views of a data set together
into one composite figure. These provide an effective approach for
exploring and digesting many different aspects of temporal data. Adding
interactivity to the ensemble can greatly enhance the exploration
process.

This paper describes new software, the \CRANpkg{tsibbletalk} package,
for exploring temporal data using linked views and time wrapping. We
first provide some background to the approach based on setting up data
structures and workflow, and give an overview of interactive systems in
R. The section following introduces the \CRANpkg{tsibbletalk} package.
We explain the mechanism for constructing interactivity, to link between
multiple hierarchical data objects and hence plots, and describe the set
up for interactively slicing and dicing time to wrap a series on itself
to investigate periodicities.

\hypertarget{background-tidy-temporal-data-and-workflow}{%
\section{Background: tidy temporal data and
workflow}\label{background-tidy-temporal-data-and-workflow}}

The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduced a
unified temporal data structure, referred to as a \texttt{tsibble}, to
represent time series and longitudinal data in a tidy format
\citep{wickham2014tidy}. A \texttt{tsibble} extends the
\texttt{data.frame} and \texttt{tibble} classes with the temporal
contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
declares a data column that holds time-related indices. The \texttt{key}
identifies a collection of related series or panels observed over the
\texttt{index}-defined period, which can comprise multiple columns. An
example of a \texttt{tsibble} can be found in the monthly Australian
retail trade turnover data (\texttt{aus\_retail}), available in the
\CRANpkg{tsibbledata} package \citep{R-tsibbledata}, shown below. The
\texttt{Month} column holds year-months as the \texttt{index}.
\texttt{State} and \texttt{Industry} are the identifiers for these 152
series, which form the \texttt{key}. Note that the column
\texttt{Series\ ID} could be an alternative option for setting up the
\texttt{key}, but \texttt{State} and \texttt{Industry} are more readable
and informative. The \texttt{index} and \texttt{key} are ``sticky''
columns to a \texttt{tsibble}, forming critical pieces for fluent
downstream temporal data analysis.

\begin{Schunk}
\begin{Soutput}
#> # A tsibble: 64,532 x 5 [1M]
#> # Key:       State, Industry [152]
#>   State                Industry                    `Series ID`    Month Turnover
#>   <chr>                <chr>                       <chr>          <mth>    <dbl>
#> 1 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Apr      4.4
#> 2 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 May      3.4
#> 3 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Jun      3.6
#> 4 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Jul      4  
#> 5 Australian Capital ~ Cafes, restaurants and cat~ A3349849A   1982 Aug      3.6
#> # ... with 64,527 more rows
\end{Soutput}
\end{Schunk}

In the spirit of tidy data from the \CRANpkg{tidyverse}
\citep{Wickham2019}, the \textbf{tidyverts} suite features
\texttt{tsibble} as the foundational data structure, and helps to build
a fluid and fluent pipeline for time series analysis. Besides
\CRANpkg{tsibble}, the \CRANpkg{feasts} \citep{R-feasts} and
\CRANpkg{fable} \citep{R-fable} packages fill the role of statistical
analysis and forecasting in the \textbf{tidyverts} ecosystem. During all
the steps of a time series analysis, the series of interest, denoted by
the \texttt{key} variable, typically persist, through the trend modeling
and also forecasting. We would typically want to examine the series
across all of the keys.

Figure \ref{fig:highlight-retail} illustrates examining temporal data
with many keys. The data has 152 series corresponding to different
industries in retail data. The multiple series are displayed using an
overlaid time series plot, along with a scatterplot of two variables
(trend versus seasonal strength) from feature space, where each series
is represented by a dot. The feature space is computed using the
\texttt{features()} function from \CRANpkg{feasts}, which summarises the
original data for each series using various statistical features. This
function along with other \textbf{tidyverts} functions is
\texttt{tsibble}-aware, and outputs a table in a reduced form where each
row corresponds to a series, which can be graphically displayed as in
Figure \ref{fig:highlight-retail-2}.

\begin{Schunk}
\begin{figure}

{\centering \subfloat[\label{fig:highlight-retail-1}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-1} }\subfloat[\label{fig:highlight-retail-2}]{\includegraphics[width=.49\linewidth]{figure/highlight-retail-2} }

}

\caption[Plots for the \code{aus\_retail} data, with the series of strongest seasonal strength highlighted]{Plots for the \code{aus\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).}\label{fig:highlight-retail}
\end{figure}
\end{Schunk}

Figure \ref{fig:highlight-retail} has also been highlighted to focus on
the one series with the strongest seasonality. To create this
highlighting, one needs to first filter the interesting series from the
features table, and join back to the original \texttt{tsibble} in order
to examine its trend in relation to others. This procedure can soon grow
cumbersome if many series are to be explored. It illustrates a need to
query interesting series on the fly. Although these two plots are
static, we can consider them as linked views because the common
\texttt{key} variables link between the two data tables producing the
two plots. This motivates the work in this package, described in this
paper, to enable interactivity of \texttt{tsibble} and
\texttt{tsibble}-derived objects for rapid exploratory data analysis.

\hypertarget{overview-of-interactivity}{%
\section{Overview of interactivity}\label{overview-of-interactivity}}

There is a long history of interactive data visualization research and
corresponding systems. Within R, the systems can be roughly divided into
systems utilizing web technology and those that do not.

R \CRANpkg{shiny} \citep{R-shiny} and \CRANpkg{htmlwidgets}
\citep{R-htmlwidgets} provide infrastructure connecting R with HTML
elements and JavaScript that support the interactivity. The
\CRANpkg{htmlwidgets} package makes it possible to embed JavaScript
libraries into R so that users are able to write only R code to generate
web-based plots. Many JavaScript charting libraries have been ported to
R as HTML widgets, including \CRANpkg{plotly} \citep{plotly2020},
\CRANpkg{rbokeh} \citep{R-rbokeh}, and \CRANpkg{leaflet}
\citep{R-leaflet} for maps. Interactions between different widgets can
be achieved with \CRANpkg{shiny} or \CRANpkg{crosstalk}
\citep{R-crosstalk}. The \CRANpkg{crosstalk} extends
\CRANpkg{htmlwidgets} with shared R6 instances to support linked
brushing and filtering across widgets, without relying on
\CRANpkg{shiny}.

Systems without the web technology include \CRANpkg{grDevices},
\CRANpkg{loon} \citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
\citep{xie_reactive_2014} based on Qt. They offer a wide array of
pre-defined interactions, such as selecting and zooming, to manipulate
plots via mouse action, keyboard strokes, and menus. The
\pkg{cranvastime} package \citep{cheng_enabling_2016} is an add-on to
\pkg{cranvas}, which provides specialized interactions for temporal
data, such as wrapping and mirroring.

The techniques implemented in the work described in this paper utilize
web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R
\CRANpkg{shiny}.

\hypertarget{using-a-shared-temporal-data-object-for-interactivity}{%
\section{Using a shared temporal data object for
interactivity}\label{using-a-shared-temporal-data-object-for-interactivity}}

The \CRANpkg{tsibbletalk} package introduces a shared tsibble instance
built on a \texttt{tsibble}. This allows for seamless communication
between different plots of temporal data. The
\texttt{as\_shared\_tsibble()} function turns a \texttt{tsibble} into a
shared instance, \texttt{SharedTsibbleData}, which is a subclass of
\texttt{SharedData} from \CRANpkg{crosstalk}. This is an R6 object
driving data transmission across multiple views, due to its mutable and
lightweight properties. The \CRANpkg{tsibbletalk} package aims to
streamline interactive exploration of temporal data, with the focus of
temporal elements and structured linking.

\hypertarget{linking-between-plots}{%
\subsection{Linking between plots}\label{linking-between-plots}}

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
categorical variable linking, where selecting one or more observations
in one category will broadcast to all other observations in this
category. That is, linking is by key variables: within the time series
plot, click on any data point, and the whole line will be highlighted in
response. The \texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s
\texttt{key} variables to achieve these types of linking.

The approach can also accommodate temporal data of nesting and crossing
structures. These time series are referred to as hierarchical and
grouped time series in the literature \citep{fpp}. The
\texttt{aus\_retail} above is an example of grouped time series. Each
series in the data corresponds to all possible combinations of the
\texttt{State} and \texttt{Industry} variables, which means they are
intrinsically crossed with each other. When one key variable is nested
within another, such as regional areas within a state, this is
considered to be a hierarchical structure.

The \texttt{spec} argument in \texttt{as\_shared\_tsibble()} provides a
means to construct hybrid linking, that incorporates hierarchical and
categorical linking. A symbolic formula can be passed to the
\texttt{spec} argument, to define the crossing and/or nesting
relationships among the key variables. Adopting \citet{Wilkinson1973}'s
notation for factorial models, the \texttt{spec} follows the \texttt{/}
and \texttt{*} operator conventions to declare nesting and crossing
variables, respectively. The \texttt{spec} for the \texttt{aus\_retail}
data is therefore specified as \texttt{State\ *\ Industry} or
\texttt{Industry\ *\ State}, which is the default for the presence of
multiple \texttt{key} variables. If there is a hierarchy in the data,
using \texttt{/} is required to indicate the parent-child relation, for
a strictly one directional \texttt{parent/child}.

To illustrate nesting and crossing we use the \texttt{tourism\_monthly}
dataset \citep{tourism} packaged in \CRANpkg{tsibbletalk}. It contains
monthly domestic overnight trips across Australia. The \texttt{key} is
comprised of three identifying variables: \texttt{State},
\texttt{Region}, and \texttt{Purpose} (of the trip), in particular
\texttt{State} nesting of \texttt{Region}, crossed together with
\texttt{Purpose}. This specification can be translated as follows:

\begin{Schunk}
\begin{Sinput}
library(tsibble)
library(tsibbletalk)
tourism_shared <- tourism_monthly %>% 
  as_shared_tsibble(spec = (State / Region) * Purpose)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{img/tourism-linking}

}

\caption[Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \code{tourism\_shared} object]{Snapshot of exploring an ensemble of linked plots of the Australian tourism data, built on a \code{tourism\_shared} object. It also illustrates persistent linked brushing to compare two groups.}\label{fig:tourism-linking-fig}
\end{figure}
\end{Schunk}

\noindent There is a three-level hierarchy: the root node is implicitly
Australia, geographically disaggregated to states, and lower-level
tourism regions. A new handy function \texttt{plotly\_key\_tree()} has
been implemented to help explore the hierarchy. It interprets
hierarchies in the shared tsibble's \texttt{spec} as a tree view, built
with \CRANpkg{plotly}. The following code line produces the linked tree
diagram (left panel of Figure \ref{fig:tourism-linking-fig}). The visual
for the tree hierarchy detangles a group of related series and provides
a bird's eye view of the data organization.

\begin{Schunk}
\begin{Sinput}
p_l <- plotly_key_tree(tourism_shared, height = 1100, width = 800)
\end{Sinput}
\end{Schunk}

The tree plot provides the graphics skeleton, upon which the rest of the
data plots can be attached. In this example, small multiples of line
plots are placed at the top right of Figure
\ref{fig:tourism-linking-fig} to explore the temporal trend across
regions by the trip purpose. The shared tsibble data can be directly
piped into \CRANpkg{ggplot2} code to create this.

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
p_tr <- tourism_shared %>%
  ggplot(aes(x = Month, y = Trips)) +
  geom_line(aes(group = Region), alpha = .5, size = .4) +
  facet_wrap(~ Purpose, scales = "free_y") +
  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
\end{Sinput}
\end{Schunk}

These line plots are heavily overplotted. To tease apart structure in
the multiple time series, the \texttt{features()} function computes
interesting characteristics, including the measures of trend and
seasonality. These are displayed in the scatterplot at the bottom right,
where one dot represents one series.

\begin{Schunk}
\begin{Sinput}
library(feasts)
tourism_feat <- tourism_shared %>%
  features(Trips, feat_stl)
p_br <- tourism_feat %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
  geom_point(aes(group = Region), alpha = .8, size = 2)
\end{Sinput}
\end{Schunk}

There is one final step, to compose the three plots into an ensemble of
coordinated views for exploration, shown in Figure
\ref{fig:tourism-linking-fig}. (This is the interactive realization of
Figure \ref{fig:highlight-retail}).

\begin{Schunk}
\begin{Sinput}
library(plotly)
subplot(p_l,
  subplot(
    ggplotly(p_tr, tooltip = "Region", width = 1100),
    ggplotly(p_br, tooltip = "Region", width = 1100),
    nrows = 2),
  widths = c(.4, .6)) %>%
  highlight(dynamic = TRUE)
\end{Sinput}
\end{Schunk}

Since all plots are created from one shared tsibble data source, they
are self-linking views. Nodes, lines, and points are hoverable and
clickable. Given the \texttt{spec}, clicking either one element in any
plot highlights all points that match the \texttt{Region} category, that
is, categorical linking. Figure \ref{fig:tourism-linking-fig} is a
static view of an interactive exploration. The steps in getting to this
point were:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A branch of the tree corresponding to Western Australia was first
  selected. (The names of the regions are a little odd, which is a quirk
  of the data set, but all four areas, Australia's South West, \ldots.,
  correspond to tourist destinations in Western Australia. Hovering over
  the node on the branch brings up the state name.) This generated the
  response in the line plots and the scatterplot that colored
  corresponding time series and points as blue.
\item
  To enable persistent selection, in oder to compare regions or states,
  ``Shift'' and click on the tree was done, after switching the color to
  red. This generated the response that points and time series
  corresponding to Sydney were highlighted in red.
\item
  Hovering over the points brings up the label for Sydney.
\end{enumerate}

Domestic tourism sees Sydney as one of the most popular destinations in
the realm of business and friends visiting over the years. Despite the
relatively weaker performance in Western Australia, Australia's North
West region sees a strongest upward trend in business, bypassing Sydney
in some years.

In summary, shared tsibble data nicely bridges between the
\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
using the common ``key''. The \texttt{as\_shared\_tsibble()} provides a
symbolic user interface for the effortless construction of a hybrid of
hierarchical and categorical linking between plots. The
\texttt{plotly\_key\_tree()} function, in turn, decodes the hierarchical
specification to plot a tree for data overview and navigation, when
accompanied by more detailed plots.

\hypertarget{slicing-and-dicing-time}{%
\subsection{Slicing and dicing time}\label{slicing-and-dicing-time}}

An important aspect of temporal data is the time context. Time has a
cyclical structure, that may correspond to seasonal patterns to be
discovered. The \texttt{index} component of the (shared) tsibble data
forms the basis for exploring seasonality. To investigate for periodic
or aperiodic patterns, series should be wrapped on themselves, where the
index is broken into temporal components like quarter or day. We shall
explore this with pedestrian traffic in Melbourne, Australia.

\begin{Schunk}
\begin{figure}

{\centering \subfloat[Initial overview state\label{fig:wrap-ped-1}]{\includegraphics[width=\textwidth]{img/wrap-0} }\newline\subfloat[1-day state\label{fig:wrap-ped-2}]{\includegraphics[width=\textwidth]{img/wrap-1} }\newline\subfloat[7-day state, anchoring to Monday\label{fig:wrap-ped-3}]{\includegraphics[width=\textwidth]{img/wrap-7} }

}

\caption[Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly]{Snapshots wrapping after slicing the \code{pedestrian20} data at different intervals, (a) none, (b) daily and (c) weekly. This type of interaction is made possible with Shiny elements.}\label{fig:wrap-ped}
\end{figure}
\end{Schunk}

The city of Melbourne has sensors installed at various locations, to
record hourly counts of pedestrians, in order to capture the daily
rhythms of the downtown \citep{ped}. Figure \ref{fig:wrap-ped} shows the
first five months of 2020 foot traffic at four different locations, for
three different time slices, daily, weekly and full five months. Plot
\ref{fig:wrap-ped-1} shows hourly counts from January to May on an
absolute timeline, facetted by locations. The stage 3 COVID-19 lockdown,
on March 16, is marked by a change of color. (The pre-lockdown period is
colored with dark green and lockdown with orange.) We can see a
significant decline in foot traffic at all four locations. QV Market is
less affected probably because this is a major produce market, an
essential service that continued to operate. Bourke St, a major shopping
center, sees a gradual uptick in the last weeks of the period indicating
that people were getting back into the shops.

Figure \ref{fig:wrap-ped-2} and \ref{fig:wrap-ped-3} show the slicing
and wrapping of the series into daily and weekly sections, respectively.
Multiple seasonalities pop out. There tends to be a daily pattern,
especially visible at the main train station, Southern Cross Station.
There is also a weekday vs weekend pattern, also most visible at
Southern Cross Station. These seasonal patterns are still present during
the lockdown, but the magnitude is greatly reduced. Numbers are also
down at the produce market and the shopping center. Birrarung Marr is
the most affected. This is the location of special events, and it is
clear that these have completely disappeared during the lockdown.

The wrapping procedure involves slicing the time index into seasonal
periods of interest, and the result is diced time. For example, hourly
pedestrian data can be decomposed into 24-hour blocks, which then
overlays the counts for all respective days, as done in plot
\ref{fig:wrap-ped-2}. For exploration, this slice position should be
controlled interactively, so that many different slices can be examined
rapidly. This can be achieved using shiny, with the functions provided
in the \CRANpkg{tsibbletalk}.

This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
reusable component that could be embedded in any shiny application. In
general, a shiny module provides a vehicle for modularising shiny
applications, relevant for both users and developers. As with all shiny
modules, the first argument in both functions in \CRANpkg{tsibbletalk}
requires a user-supplied id string that must be unique. The UI function
\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
the number of periods to be diced. The workhorse is the server function
\texttt{tsibbleWrapServer()}, encapsulating the algorithm that
transforms data and sends messages to update the plot accordingly. The
\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
object, where one can plot data using either lines or other graphical
elements (such as boxplots). As the function name suggests, a (shared)
tsibble is needed to start the engine, so that the time \texttt{index}
can be retrieved for dissection. The \texttt{period} option semantically
takes a desired number of seasonal periods to be shifted, for example
data shifted by ``1 day'', ``2 days'', or ``1 week'', etc. In other
words, the \texttt{period} defines the grind level. For date-times
(represented by \texttt{POSIXt}), the granularity ranges from fine
``day'' to a much coarser ``year''. The following code snippet generates
Figure \ref{fig:wrap-ped}. The creation of the \texttt{pedestrian20}
data is available in supplementary R files.

\begin{Schunk}
\begin{Sinput}
library(shiny)
p_line <- pedestrian20 %>%
  ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
  geom_line(size = .3) +
  facet_wrap(~ Sensor, scales = "free_y") +
  labs(x = "Date Time") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")

ui <- fluidPage(
  tsibbleWrapUI("dice")
)
server <- function(input, output, session) {
  tsibbleWrapServer("dice", ggplotly(p_line, height = 700), period = "1 day")
}
shinyApp(ui, server)
\end{Sinput}
\end{Schunk}

Figure \ref{fig:wrap-ped-1} corresponds to the initial state, with the
slider incremented by 1-day units. The ``play'' button near the end of
the slider can automatically animate the slicing and dicing process,
walking the viewer through all 24 hours of the 152 days. Alternatively,
users can drag the slider to examine selected slices.

In response to the slider input, the plot will be updated and loaded
with newly transformed data. At its core, keeping the application as
performant as possible is the top priority. Without completely redrawing
the plot, the \texttt{plotlyProxy()} react method is invoked internally
for talking to shiny. The underlying tsibble data is being called back
and processed in R. Only transformed data gets fed back to the shiny
server, for updating with resetting the x-axis ranges and breaks. The
other plot configurations, such as marks, y-axes, and layouts, are
cached and used as is.

The new shiny module exploits the temporal aspect for a tsibble object,
available through the \texttt{index} attribute. It allows users to slide
through relative periods to digest seasonal behaviors, with a nimble
user experience.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

At the heart of the \CRANpkg{tsibbletalk} package is a blending of the
best bits from \CRANpkg{tsibble}, \CRANpkg{crosstalk}, \CRANpkg{plotly},
and \CRANpkg{shiny}.

The \texttt{as\_shared\_tsibble()} turns a \texttt{tsibble} object to a
shared data class, with an option to express any nesting and crossing
structures from the \texttt{key} attribute. If nesting is found in the
data, the \texttt{plotly\_key\_tree()} creates an interactive
hierarchical tree to help with the data overview. This sets the stage
for hierarchical and categorical linking between multiple views from one
shared tsibble.

A new shiny module, \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, provides a lens for looking at temporal
aspects of a \texttt{tsibble}, in particular seasonal or cyclical
variations. The slicing and dicing technique efficiently wrap time lines
for user-defined plots. The \texttt{plotlyProxy()} react method makes it
possible to send wrapped data to the server and amend the plot straight
way.

\bibliography{tsibbletalk.bib}

\address{%
Earo Wang\\
The University of Auckland\\%
Department of Statistics\\
%
%
%
\href{mailto:earo.wang@auckland.ac.nz}{\nolinkurl{earo.wang@auckland.ac.nz}}%
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
%
%
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
}